Classifying galaxy spectra at $0.5<z<1$ with self-organizing maps

### 2018-03-25 sections0.10/method0.10.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

@@

\changed{In this section we  briefly review the use of both self-organizing maps and K-means clustering in astronomy and describe their algorithms.} 
As K-means is both simpler and more well-known, we introduce it briefly first before discussing self-organizing maps in detail.

@@

\changed{
K-means clustering~\citep{Macqueen67} is an unsupervised method used in many astronomical studies~\citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-means clustering divides data into $K$ clusters, in such a way that each data point belongs to the cluster with the nearest mean value.
To first order, SOM and K-means work similarly: both algorithms operate on the distances between \boldit{N} $\in \Re^m$ objects defined by the properties under consideration.
The SOM algorithm also considers distances between objects in the low-dimensional map, as controlled by a neighborhood function. 
When the radius of the neighbourhood function goes to zero, the SOM algorithm loses its ordering power and acts like K-means clustering.
}
%Referee: The objective function for k-means (equation 1, page 4) is not quite clear as it does not indicate what the "min" is taken over; I suggest adding a subscript "k" here (assuming that is what was intended).
% DONE

The K-means method partitions a data set $x_i$, having $n$ points in $d$-dimensional feature space, into $K$ clusters.
The number of clusters $K\leq n$ must be decided by the user in advance, and the cluster centroids in feature space are located at positions $\mu_k$.
The goal of the algorithm is to minimize the sum
\begin{equation}
\label{eq:km_sos}
J = \sum_{i=1}^n \sum_{k=1}^K {\min}_k \big( \norm{x_i - \mu_k}^2 \big)
\end{equation}
where $\norm{z}$ indicates the distance measurement in $d$-dimensional space. In this work the Euclidean distance is is used, but other metrics are also possible.
\changed{The algorithm is initialized by randomly choosing $K$ points as cluster centroids.}
Each data point is assigned to the cluster with the closest centroid, and the centroids are re-calculated.
The steps of re-assigning cluster membership and re-calculating the cluster centroids are repeated until a stopping criterion is reached (a maximum number of iterations or no change in cluster assignments).
We used the \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform  K-means clustering.
The default maximum number of iterations (100) was used.

@@

\changed{ 
 Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16a,Hossein16b,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

@@

Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can distinguish an astronomical object from instrumental artifact, and then classify the object based on its spectrum.
}

@@

     We use several metrics to measure the degree of agreement between classification methods. 
     As a baseline method for assigning galaxy spectra to template classes, we compute the value of chi-squared between each galaxy and each template and assign the galaxy to the template class with lowest chi-squared value. 
     A metric of agreement with this classification is an agreement score, computed as follows.
    If a method sorted a galaxy into the same cluster or neuron as its best-fit chi-square template, this was considered a match and scored 1 point. 
    For the larger-sized self-organized maps, if a galaxy was classified between two nodes, one of which contained its best-fit chi-square template, this scored 0.5 point. 
    Otherwise, the classification score was zero points.     
     Another method to check the degree of agreement between different methods is the Fleiss kappa index \citet{landis77}. 
    This index measures agreement between three or more raters that are classifying a specific number of objects to a number of classes and was calculated using the R software package.

>>>>

%Referee comments
%1d. The authors stated "We have added a footnote to clarify how chi-squared was computed."  I cannot find this footnote.  Please define this method in section 3.

%1e. The metrics proposed in section 3.4 are
%- agreement of algorithm X's output with that obtained by a chi-squared minimization
%- Fleiss kappa
%- silhouette score (S) - also referred to as within-cluster similarity

%I suggest referring to the first metric as "chi-squared agreement" (here and throughout, including the discussion of results starting in section 4.3) or some other term to capture the fact that it specifically relates an algorithm to the chi-squared results.

%I am still confused by the the Fleiss kappa results.  If the chi-squared method is the point of reference to use (as with the proposed agreement metric), why does it not appear here?  What does it tell us to determine that the k-means, supervised ANN, and SOMs are in "fair" agreement? (with each other?)  Is that a goal?  The chi-squared agreement analysis showed higher agreement with the SOM than with the other methods, so I wouldn't expect (nor want) there to be high kappa values between those three.  As currently presented, I don't think this adds anything to the paper, and I suggest omitting it (or rephrasing as a second measure of agreement with chi-squared, and computing values accordingly).

%2. The re-organization of the paper is helpful for the Introduction, but the new content needs to be better integrated into sections 3.1 and 3.2.  Currently it reads disjointedly.

%2a. Section 3.1: The notation "N \in R^m" does not make sense, especially since m is not defined.  Consider something like "N objects, each represented by m features" or "N objects x \in R^m, where m is the number of features used to represent each object".

%Note that the following paragraph uses a lower-case "n" for the number of objects.  Please make these consistent.

%2b. Section 3.2:
%- The text prior to 3.2.1 is now rather unwieldy and hard to follow.  It jumps back and forth between ANNs and SOMs. I suggest making a new 3.2 that covers "Artificial neural networks" (since the paper treats them as an alternative to SOMs later) and 3.3 can be devoted to SOMs as a specialized (semi-supervised) version of ANNs.  Merge old and new content to make a logical flow.
%- Provide a citation for ANNs when first mentioned.
%- What is meant by "a set of training methods" (end of page 3, column 2)?  Typically only a single training method (e.g., backpropagation) is used.
%- Standard practice is to avoid having a single subsection (3.2.1). Remove the subsection heading, or make two (3.2.1, 3.2.2) on different sub-topics.  For example, section 3.3 could (should) be merged into the SOM section, as a subsection.

% from minor ref comments:
%2. The paper now asserts that ANNs outperform chi-square methods (top of page 4, left column) - in what sense?  Accuracy (which as we've discussed may not be measurable)?  Runtime?  Something else?

@@

In this section we  briefly review the use of both self-organizing maps and K-means clustering in astronomy and describe their algorithms. 
As K-means is both simpler and more well-known, we introduce it briefly first before discussing self-organizing maps in detail.

@@

K-means clustering~\citep{Macqueen67} is an unsupervised method used in many astronomical studies~\citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-means clustering divides data into $K$ clusters, in such a way that each data point belongs to the cluster with the nearest mean value.
To first order, SOM and K-means work similarly: both algorithms operate on the distances between \boldit{N} $\in \Re^m$ objects defined by the properties under consideration.
The SOM algorithm also considers distances between objects in the low-dimensional map, as controlled by a neighborhood function. 
When the radius of the neighbourhood function goes to zero, the SOM algorithm loses its ordering power and acts like K-means clustering.

The K-means method partitions a data set $x_i$, having $n$ points in $d$-dimensional feature space, into $K$ clusters.
The number of clusters $K\leq n$ must be decided by the user in advance, and the cluster centroids in feature space are located at positions $\mu_k$.
The goal of the algorithm is to minimize the sum
\begin{equation}
\label{eq:km_sos}
J = \sum_{i=1}^n \sum_{k=1}^K {\min}_k \big( \norm{x_i - \mu_k}^2 \big)
\end{equation}
where $\norm{z}$ indicates the distance measurement in $d$-dimensional space. In this work the Euclidean distance is is used, but other metrics are also possible.
The algorithm is initialized by randomly choosing $K$ points as cluster centroids.
Each data point is assigned to the cluster with the closest centroid, and the centroids are re-calculated.
The steps of re-assigning cluster membership and re-calculating the cluster centroids are repeated until a stopping criterion is reached (a maximum number of iterations or no change in cluster assignments).
We used the \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform  K-means clustering.
The default maximum number of iterations (100) was used.

@@

 Artificial neural networks (ANNs), which are inspired by the way neurons in a human brain route and process data, are very powerful tools that are used in data processing and pattern recognition problems.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about nonlinear and complex relations between input and output data, and how to apply these relations to new sets of data~\citep[e.g.][]{Hossein14,Hossein16a,Hossein16b,Ellison16a, Ellison16b}.
Studies have shown that ANNs outperform chi-square minimizing techniques and can be used as an alternative choice for fitting data~\citep[e.g.][]{Marquez91}.
Specifically, ANNs perform faster in large databases~\citep[][]{Gulati97}.

@@

Large spectroscopic surveys have made available integrated spectra of millions of galaxies.
These integrated spectra combine the light of billions of individual stars and nebulae within a galaxy, and
finding patterns and common characteristics between galaxies can be a complex task.
\citet{In12} introduced a new clustering tool based on the SOM method for analyzing these large datasets.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09} to test their tool, and created very large SOMs to analyze the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of spectra. 
Later, \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys to find a new class of quasars.
The other application of SOMs is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from the GAIA survey that were previously classified as ``unknown'' by the SDSS pipeline. This package can distinguish an astronomical object from instrumental artifact, and then classify the object based on its spectrum.

@@

     We use several metrics to measure the degree of agreement between classification methods. 
     As a baseline method for assigning galaxy spectra to template classes, we compute the value of chi-squared between each galaxy and each template%
\footnote{The \citetalias{Kinney96} template spectra do not have tabulated uncertainties and we assume them to be photon-noise dominated for the purposes of computing $\chi^2$.}
     and assign the galaxy to the template class with lowest chi-squared value. 
     A metric of agreement with this classification is an agreement score, computed as follows.
    If a method sorted a galaxy into the same cluster or neuron as its best-fit chi-square template, this was considered a match and scored 1 point. 
    For the larger-sized self-organized maps, if a galaxy was classified between two nodes, one of which contained its best-fit chi-square template, this scored 0.5 point. 
    Otherwise, the classification score was zero points.     
     Another method to check the degree of agreement between different methods is the Fleiss kappa index \citet{landis77}. 
    This index measures agreement between three or more raters that are classifying a specific number of objects to a number of classes and was calculated using the R software package.

### 2018-03-25 sections0.10/results0.10.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

@@

%Referee: Good analysis in section 4.1.2.  Since the cluster labels are arbitrary ("cluster 1" for k-means does not mean the same thing as "cluster 1" for SOM), I suggest using letters for one algorithm (cluster A, B, C...) to make this very obvious.  Or, call the clusters K1, K1, K3... for k-means and S1, S2, S3... for SOM.

%referee: K-means cluster 4, which has only one item in it (Sc), genuinely seems to be different given its peak between 3000-4000 A.  Placing it in SOM cluster 3 obscures this feature.  This is worth mentioning in the discussion.

@@

    \begin{figure*}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[width=.99\textwidth, height=         7.5cm]{k_means_images/classified_group_in_4cluster_kinney.png}
    \end{subfigure}
    \hfill
        \begin{subfigure}[b]{0.49\textwidth}
        \centering \includegraphics[width=.99\textwidth, height=         7.5cm]{k_means_images/classified_group_in_4cluster_som.png}
    \end{subfigure}
    \caption{Clustering the \citetalias{Kinney96} template spectra using K-means clustering     (left) and SOM (right). Note that both methods randomly assign the initial values for their analysis, therefore the cluster number is only used for distinguishing cluster membership.}
    \label{fig: som_k_means_4}
    \end{figure*}

    We performed K-means clustering on the \citetalias{Kinney96} template spectra using 4 clusters.

    Fig.~\ref{fig: som_k_means_4} shows the clustering of the template spectra into 4 groups using both K-means and SOMs.
    Both methods separated the most starburst-like and quiescent galaxies in a similar manner.
    However, they show discrepancies for spectral types between these extreme cases.
    Table~\ref{tab:kmeans_som4} shows the group assignments of the \citetalias{Kinney96} templates in the two methods.
    Templates B, E, S0 and Sa are grouped in the same cluster/neuron in both K-means and $1\times4$-sized SOM. 
    K-means grouped the Sb template with the above templates, but the SOM network classified it as a separate group.
    The K-means method clustered the Sc template as a separate group, while the SOM network grouped it with SB5 and SB6. 
    Another noticeable difference is that K-means clustered the SB1 and SB2 templates in one group and SB3 to SB6 in another while
    the SOM network grouped the SB1 to SB4 templates in one neuron and SB5 and SB6 in another.
    As Fig.~\ref{fig: som_k_means_4} shows, K-means clustering of the templates produced groups more similar to each other, while SOM emphasizes the templates' differences.

    \begin{table}
    \centering
    \begin{tabular}{||l|c|c||}
    \hline
    \hline
    \citetalias{Kinney96} template & K-means group & SOM group \\
    \hline
    B                                & 3               & 1                         \\
    E                                & 3               & 1                         \\
    S0                               & 3               & 1                         \\
    Sa                               & 3               & 1                         \\
    Sb                               & 3               & 2                         \\
    Sc                               & 4               & 3                         \\
    SB1                              & 2               & 4                         \\
    SB2                              & 2               & 4                         \\
    SB3                              & 1               & 4                         \\
    SB4                              & 1               & 4                         \\
    SB5                              & 1               & 3                         \\
    SB6                              & 1               & 3                        \\
    \hline
    \end{tabular}
    \caption{Group assignments for \citetalias{Kinney96} template galaxies, using K-means and $1\times4$-sized SOM methods}
    \label{tab:kmeans_som4}
    \end{table}

@@

            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have spectral types similar to those of the quiescent galaxies and the spectral types of the rest are similar to the starburst galaxies.
            The lower left panel shows the median spectrum of the 80 galaxies that were classified as quiescent. 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the spectra clearly show the 4000\AA~break, one of the signatures of quiescent galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar spectral types to Sa galaxies, but with stronger emission lines.
            The median spectrum in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.45\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/SED_total1by2_fluxdensity.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption{Classification of fitted galaxy spectra from \citet{Hossein12} using the $1\times2$~network trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 spectra are classified into the quiescent group while the 62 of the 142 spectra are categorized as starburst galaxies. Lower panel: median spectra of the galaxies in each group (flux densities are normalized by dividing spectra by flux density at 4000\AA), quiescent on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

@@

 %referee: How do we conclude from Fig 5 that "As Fig. 5 shows, K-means clustering of the templates produced groups more similar to each other, while SOM emphasizes the templates’ differences"?  To my eye, SOM clusters 1 and 2 are very similar to each other, and more so than any pair of k-means clusters.  To support the claims about cluster similarity, one can measure this empirically with, e.g., the pairwise Euclidean distance between cluster (means).  Please update the end of section 4.3 as necessary after performing this analysis.

@@

       To compare results from SOM, K-means, and unsupervised ANN, we first consider how well these methods matched a simple chi-squared fitting classification. 
        Matching galaxy spectra to templates using chi-squared, we found that not all of the templates had matches among the galaxy sample: no galaxy spectra were best-matched to the B, E, or Sc templates. 
        This is consistent with the results from the self-organized maps.
        Summing the classification score over all galaxies in the sample allows us to compare how well each technique matched the results of chi-squared fitting.
        The totals were K-means 124, $1\times4$-sized SOM 103, $1\times12$-sized SOM 111.5, $1\times22$-sized SOM 79, and unsupervised ANN 52.
        This shows that the K-means method and $1\times12$-sized SOM have the most similarity with minimum chi-square fitting.
        Summing the classification score over all methods for a single galaxy shows how well the methods agree for a particular spectrum.
        For example, if the total score of a galaxy is 5, it means that all techniques classified the galaxy into the same group.
        Comparing the scores, we found that 12 per cent of galaxies scored 5, 36 per cent of galaxies had a score greater than 4, and 56 per cent had a total score of 3 or more.
        The score-5 galaxies classified similarly by all methods were primarily (60 per cent) in the Sa template group. 
        The score $>=3$ galaxies were predominantly SB1, Sa, and Sb type (32, 28, 18 per cent, respectively).
        The different methods show the most agreement for galaxies similar to the SB1 and Sa templates, with less agreement on the high-extinction starburst types (SB5, SB6).

           %referee: On page 12, I am confused by the Fleiss kappa results.  Since it allows pairwise comparison of two raters, I was anticipating perhaps an upper triangular matrix of all (5 choose 2) pairs of methods compared to each other.  Instead, only two values are reported (27% and 21%) and it is not clear what (which pairs) they characterize.  Please elaborate.

       We also used the Fleiss kappa index \citet{landis77} to 
        measure the agreement between methods. 
        Here the `raters' are the different methods classifying a specific number of objects (i.e. 142 galaxies) to a number of classes (i.e. 12 templates).
        We computed two different values of Fleiss kappa, between  K-means, unsupervised ANN, and either 
         $1\times12$-sized or $1\times22$-sized SOMs.
        The resulting values were $\sim 27$ per cent and $\sim 21$ per cent, respectively.
        According to \citet{landis77}, the interpretation of a Fleiss kappa measure in the range 21--40 per cent is that the raters are in fair agreement, consistent with the results above.

A final comparison between the classification methods is 
made with the silhouette score  \citep{rousseeuw87}.
The highest score among our classification was for K-means (0.549), followed by $1\times4$-sized SOM (0.371), $1\times12$-sized SOM (0.302) and $1\times22$-sized SOM (0.249).
The chi-squared fitting resulted in a score of 0.221 and the supervised neural network 0.142.
As discussed in Section~\ref{sec:metrics}, the silhouette score usually declines as the number of groups increases, so we cannot conclude from these values that K-means is the superior method. 
However, the $1\times12$-sized SOM, chi-squared fitting, and supervised neural network all have either 8 or 9 classes.
In this case we can compare scores directly and conclude that the self-organized map yields groups with more internal similarity.
This is in contrast to the comparison between K-means and self-organized maps for the templates alone (see Section\ref{sec: Kmeansvssom}) where K-means clustering  produced groups more similar to each other, and emphasizes that the exact data used in a clustering process can strongly affect the outcome.

@@

        \begin{figure*}
\includegraphics[width=\textwidth]{images0.01/1d/fig11_new_bw.png}
        \caption{
        From top to bottom: relations between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. 
        Circles, pentagons, and squares show the median values of these properties of galaxies in the $1\times22$, $1\times 12$ and $1\times4$ SOM networks; 
        triangles show the median value of properties for each K-means cluster (using $K=4$) and stars show median values for groups from \citetalias{Hossein12}.
        Error bars show standard deviations of the group properties.
        Colours show the galaxy types in the SOM classification. 
        The lightest grey (most pure red in the online version) colour shows E type galaxies and the black (most blue in the online version) indicates the SB1 galaxies. 
        The shading of other colours between light grey and black (red and blue in the online version) measures the location of galaxies along the SOM from  quiescent to starburst.}
        \label{fig: props_vs_props}
        \end{figure*}

>>>>

% referee comments (minor)
%3. Section 4.3: I suggest putting the numeric results for the chi-squared agreement score (paragraph 2) and silhouette score (paragraph 4) into tables to improve their readability.

%4. Section 4.2.1 - typo - space needed between the symbol for "angstroms" and the word following it (e.g., "range", "break" - in the new bolded text).

@@

@@

    \begin{figure*}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \includegraphics[width=.99\textwidth, height=         7.5cm]{k_means_images/classified_group_in_4cluster_kinney_1.pdf}
    \end{subfigure}
    \hfill
        \begin{subfigure}[b]{0.49\textwidth}
        \centering \includegraphics[width=.99\textwidth, height=         7.5cm]{k_means_images/classified_group_in_4cluster_som.png}
    \end{subfigure}
    \caption{Clustering the \citetalias{Kinney96} template spectra using K-means clustering (left) and SOM (right). Note that both methods randomly assign the initial values for their analysis, therefore the cluster labels are only used for distinguishing cluster membership.}
    \label{fig: som_k_means_4}
    \end{figure*}

    We performed K-means clustering on the \citetalias{Kinney96} template spectra using 4 clusters.
    Fig.~\ref{fig: som_k_means_4} shows the clustering of the template spectra into 4 groups using both K-means and SOMs.
    Both methods separated the most starburst-like and quiescent galaxies in a similar manner.
    However, they show discrepancies for spectral types between these extreme cases.
    Table~\ref{tab:kmeans_som4} shows the group assignments of the \citetalias{Kinney96} templates in the two methods.
    Templates B, E, S0 and Sa are grouped in the same cluster/neuron in both K-means and $1\times4$-sized SOM. 
    K-means grouped the Sb template with the above templates, but the SOM network classified it as a separate group, possibly because it is much more noisy in the ultraviolet range.
    The K-means method clustered the Sc template as a separate group, 
    likely due to its distinct features in the 3000--4000 \AA range, while the SOM network grouped it with SB5 and SB6. 
    Classifying the Sc template in the group with SB5 and SB6 obscures the peak feature in the Sc template.
    However, when the number of SOM nodes is increased (e.g. in Fig.~\ref{fig: 1by22T}), the Sc template separates and occupies a single neuron.
    Another noticeable difference is that K-means clustered the SB1 and SB2 templates in one group and SB3 to SB6 in another, while the SOM network grouped the SB1 to SB4 templates in one neuron and SB5 and SB6 in another.
    SB1 and S2 have similar strong emission lines in the blue, while SB3--6 show H$\alpha$ emission but only weaker emission lines at shorter wavelengths;
    the SB5 and SB6 templates are flatter with a less-pronounced 4000 \AA break than the remaining SB templates.
    By comparing the membership of different clusters in Table~\ref{tab:kmeans_som4} we can see that K-means clustering emphasizes the input templates' similarities, while SOM emphasizes the templates' differences.

    \begin{table}
    \centering
    \begin{tabular}{||l|c|c||}
    \hline
    \hline
    \citetalias{Kinney96} template & K-means group & SOM group \\
    \hline
    B                                & C               & 1                         \\
    E                                & C               & 1                         \\
    S0                               & C               & 1                         \\
    Sa                               & C               & 1                         \\
    Sb                               & C               & 2                         \\
    Sc                               & D               & 3                         \\
    SB1                              & B               & 4                         \\
    SB2                              & B               & 4                         \\
    SB3                              & A               & 4                         \\
    SB4                              & A               & 4                         \\
    SB5                              & A               & 3                         \\
    SB6                              & A               & 3                        \\
    \hline
    \end{tabular}
    \caption{Group assignments for \citetalias{Kinney96} template galaxies, using K-means and $1\times4$-sized SOM methods}
    \label{tab:kmeans_som4}
    \end{table}

@@

            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have spectral types similar to those of the quiescent galaxies and the spectral types of the rest are similar to the starburst galaxies.
            The lower left panel shows the median spectrum of the 80 galaxies that were classified as quiescent. 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the spectra clearly show the 4000\AA~break, one of the signatures of quiescent galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar spectral types to Sa galaxies, but with stronger emission lines.
            The median spectrum in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{0.45\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/hit_v_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.45\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/SED_total1by2_fluxdensity.png}
                \end{subfigure}
                \caption{Classification of fitted galaxy spectra from \citet{Hossein12} using the $1\times2$~network trained from the \citet{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 spectra are classified into the quiescent group while the 62 of the 142 spectra are categorized as starburst galaxies. Lower panel: median spectra of the galaxies in each group (flux densities are normalized by dividing spectra by flux density at 4000\AA), quiescent on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

@@

@@

       To compare results from SOM, K-means, and supervised ANN, we first consider how well these methods matched a simple chi-squared fitting classification. 
        Matching galaxy spectra to templates using chi-squared, we found that not all of the templates had matches among the galaxy sample: no galaxy spectra were best-matched to the B, E, or Sc templates. 
        This is consistent with the results from the self-organized maps.
        Summing the classification score over all galaxies in the sample allows us to compare how well each technique matched the results of chi-squared fitting.
        The totals were K-means 124, $1\times4$-sized SOM 103, $1\times12$-sized SOM 111.5, $1\times22$-sized SOM 79, and supervised ANN 52.
        This shows that the K-means method and $1\times12$-sized SOM have the most similarity with minimum chi-square fitting.
        Summing the classification score over all methods for a single galaxy shows how well the methods agree for a particular spectrum.
        For example, if the total score of a galaxy is 5, it means that all techniques classified the galaxy into the same group.
        Comparing the scores, we found that 12 per cent of galaxies scored 5, 36 per cent of galaxies had a score greater than 4, and 56 per cent had a total score of 3 or more.
        The score-5 galaxies classified similarly by all methods were primarily (60 per cent) in the Sa template group. 
        The score $\geq 3$ galaxies were predominantly SB1, Sa, and Sb type (32, 28, 18 per cent, respectively).
        The different methods show the most agreement for galaxies similar to the SB1 and Sa templates, with less agreement on the high-extinction starburst types (SB5, SB6).

       We also used the Fleiss kappa index \citep{Fleiss71} to 
        measure the agreement between methods. 
        Here the `raters' are the different methods classifying a specific number of objects (i.e. 142 galaxies) to a number of classes (i.e. 12 templates).
        We computed values of Fleiss kappa twice: firstly to compare results from K-means, supervised ANN, and $1\times12$-sized SOM and secondly to compare results from K-means, supervised ANN, and $1\times22$-sized SOM.
        The resulting values were $\sim 27$ per cent and $\sim 21$ per cent, respectively.
        According to \citet{landis77}, the interpretation of a Fleiss kappa measure in the range 21--40 per cent is that the raters are in fair agreement, consistent with the results above.

A final comparison between the classification methods is 
made with the silhouette score  \citep{rousseeuw87}.
The highest score among our classification was for K-means (0.549), followed by $1\times4$-sized SOM (0.371), $1\times12$-sized SOM (0.302) and $1\times22$-sized SOM (0.249).
The chi-squared fitting resulted in a score of 0.221 and the supervised neural network 0.142.
As discussed in Section~\ref{sec:metrics}, the silhouette score usually declines as the number of groups increases, so we cannot conclude from these values that K-means is the superior method. 
However, the $1\times12$-sized SOM, chi-squared fitting, and supervised neural network all have either 8 or 9 classes.
In this case we can compare scores directly and conclude that the self-organized map yields groups with more internal similarity.
This is in contrast to the comparison between K-means and self-organized maps for the templates alone (see Section~\ref{sec: Kmeansvssom}) where K-means clustering  produced groups more similar to each other, and emphasizes that the exact data used in a clustering process can strongly affect the outcome.

@@

        \begin{figure*}
\includegraphics[width=\textwidth]{images0.01/1d/fig11_new_col.png}
        \caption{
        From top to bottom: relations between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. 
        Circles, pentagons, and squares show the median values of these properties of galaxies in the $1\times22$, $1\times 12$ and $1\times4$ SOM networks; 
        triangles show the median value of properties for each K-means cluster (using $K=4$) and stars show median values for groups from \citetalias{Hossein12}.
        Error bars show standard deviations of the group properties.
        Colours show the galaxy types in the SOM classification. 
        The lightest grey (most pure red in the online version) colour shows E type galaxies and the black (most blue in the online version) indicates the SB1 galaxies. 
        The shading of other colours between light grey and black (red and blue in the online version) measures the location of galaxies along the SOM from  quiescent to starburst.}
        \label{fig: props_vs_props}
        \end{figure*}

### 2018-03-25 sections0.10/summary0.10.tex
Made new version to address latest (Feb 2018) ref report. Incorporated ref rept into tex files as comments.

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
 The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by   \citetalias{Hossein12}    could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.
    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Compared to classification using chi-squared minimization or the supervised neural network, the galaxies classed together by the self-organizing map had more similar spectra.

>>>>

    A sample of 142 high-redshift galaxy spectra from \citetalias{Hossein12} was classified by the trained networks.
  This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates although many of the spectra fall in between template classes.
     In the two-dimensional network, only 23 galaxies occupied exactly the same neurons as the \citetalias{Kinney96} template spectra (in one-dimensional networks this number was 56).
    The freedom of having in-between types is one of the main differences between supervised and unsupervised artificial neural networks: the supervised training method used by \citetalias{Hossein12} could not classify 37 out of 142 spectra in galaxy sample as matching one of the same set of templates.

    Comparing different classification methods, we found that spectra classified similarly by all methods were well-matched to quiescent templates (\citetalias{Kinney96} Sa and Sb). 
    Spectra with emission lines or ultraviolet emission showed more variable classification results.
    Classification produced by self-organized maps showed closer correspondence than the supervised method results to the classification from a chi-squared match.
    Using the Fleiss kappa index to compare classifications showed that the K-means, supervised ANN and one-dimensional SOM are in fair agreement.
    One-dimensional self-organized-map based classification resulted in a 
    higher silhouette score for than classification using chi-squared minimization or the supervised neural network, indicating that the galaxies classed together by the SOM had more similar spectra.
